---
title: "Project code"
author: "Jessica Hin (008515095)"
date: "2023-06-04"
output:
  html_document:
    df_print: paged
---

```{r setup}
setwd('/Users/jessh/ADS503')
data <- read.csv('risk_factors_cervical_cancer.csv')

library(MASS)
library(dplyr)
library(caret)
library(ggplot2)
library(reshape2)
```

```{r preprocessing pt 1}
# checking for zero variance predictors
# Hinselmann test shows up in the results, but the other three tests do not (Schiller,Cytology,Biopsy)
data_nearzero <- data[,as.integer(nearZeroVar(data))]

# remove IUD column
#data <- data[,IUD] 


# data frame with stds included
data_wstd <- data

# checking for missing values
data_miss <- sapply(data, function(x) sum(is.na(x)))


# subbing in the values with question marks as NAs
data[data == "?"] <- NA

# checking for missing values again
data_miss <- sapply(data, function(x) sum(is.na(x)))

# removing the std time cols
data <- subset(data, select = -c(STDs..Time.since.first.diagnosis, STDs..Time.since.last.diagnosis))
sapply(data, function(x) sum(is.na(x)))

# changing character variables to integers
data[,2:27] <- sapply(data[,2:27], as.integer)

# Changing response to factors

#data[,c(5,8,10,12,14:25,27:34)][data[,c(5,8,10,12,14:25,27:34)] == 0]  <- "Negative"
#data[,c(5,8,10,12,14:25,27:34)][data[,c(5,8,10,12,14:25,27:34)] == 1] <- "Positive"
data[,c(31:34)][data[,c(31:34)] == 0] <- "Negative"
data[,c(31:34)][data[,c(31:34)] == 1] <- "Positive"
data$Hinselmann <- as.factor(data$Hinselmann)
data$Schiller <- as.factor(data$Schiller)
data$Citology <- as.factor(data$Citology)
data$Biopsy <- as.factor(data$Biopsy)
```

```{r preprocessing imputation}
# creating a mode function
Mode <- function(x, na.rm = FALSE) {
  if(na.rm){
    x = x[!is.na(x)]
  }

  ux <- unique(x)
  return(ux[which.max(tabulate(match(x, ux)))])
}

# changing the hormonal and IUD into one variable, birth control
data$Birth.control <- as.integer(as.integer(data$Hormonal.Contraceptives) | as.integer(data$IUD))

# removing the IUD and hormonal contraceptive variable
data = subset(data, select = -c(Hormonal.Contraceptives, IUD, STDs.cervical.condylomatosis,STDs.vaginal.condylomatosis,STDs.vulvo.perineal.condylomatosis, STDs.AIDS))

# imputing the missing data
data$Number.of.sexual.partners[is.na(data$Number.of.sexual.partners)] <- round(mean(data$Number.of.sexual.partners, na.rm = TRUE))
data$First.sexual.intercourse[is.na(data$First.sexual.intercourse)] <- round(mean(data$First.sexual.intercourse, na.rm = TRUE))
data$Num.of.pregnancies[is.na(data$Num.of.pregnancies)] <- round(mean(data$Num.of.pregnancies, na.rm = TRUE))
data$Smokes[is.na(data$Smokes)] <- Mode(data$Smokes, na.rm = TRUE)
data$Number.of.sexual.partners[is.na(data$Number.of.sexual.partners)] <- Mode(data$Number.of.sexual.partners, na.rm = TRUE)
data$First.sexual.intercourse[is.na(data$First.sexual.intercourse)] <- round(mean(data$First.sexual.intercourse, na.rm = TRUE))
data$Num.of.pregnancies[is.na(data$Num.of.pregnancies)] <- Mode(data$Num.of.pregnancies, na.rm = TRUE)
data$Smokes[is.na(data$Smokes)] <- Mode(data$Smokes, na.rm = TRUE)
data$Smokes..years.[is.na(data$Smokes..years.)] <- round(mean(data$Smokes..years., na.rm = TRUE))
data$Smokes..packs.year.[is.na(data$Smokes..packs.year.)] <- round(mean(data$Smokes..packs.year., na.rm = TRUE))
data$Hormonal.Contraceptives..years.[is.na(data$Hormonal.Contraceptives..years.)] <- round(mean(data$Hormonal.Contraceptives..years., na.rm = TRUE))
data$IUD..years.[is.na(data$IUD..years.)] <- round(mean(data$IUD..years., na.rm = TRUE))
data$STDs[is.na(data$STDs)] <- Mode(data$STDs, na.rm = TRUE)
data$STDs..number.[is.na(data$STDs..number.)] <- Mode(data$STDs..number., na.rm = TRUE)
data[,12:20][is.na(data[,12:20])] <- Mode(data$STDs, na.rm = TRUE)
data$STDs[is.na(data$STDs)] <- Mode(data$STDs, na.rm = TRUE)
data$Birth.control[is.na(data$Birth.control)] <- Mode(data$Birth.control)

# changing the boolean variables to factors
data[,c(5,10,12:19,21:29)] <- lapply(data[,c(5,10,12:19,21:29)], as.factor)
```

Using the equation $x = \frac{p(records)-rare}{1-p}$ where p is .20 to rebalance the data set per response variable.

```{r rebalancing Hinselmann}
p = .20
records = as.integer(dim(data)[1])
rare = table(data$Hinselmann)[2]
resamp_times <- (p*records - rare)/(1-p)


pos_hins <- which(data$Hinselmann == "Positive")
resamp <- sample(pos_hins, size = resamp_times, replace = TRUE)
resamp <- data[resamp,]

rebal_hins_data <- rbind(data,resamp)
table(rebal_hins_data$Hinselmann)[2]
table(rebal_hins_data$Hinselmann)[2]/dim(rebal_hins_data)[1]
```

```{r rebalancing Schiller}
p = .20
records = as.integer(dim(data)[1])
rare = table(data$Schiller)[2]
resamp_times <- (p*records - rare)/(1-p)


pos_schill <- which(data$Schiller == "Positive")
resamp <- sample(pos_schill, size = resamp_times, replace = TRUE)
resamp <- data[resamp,]

rebal_schill_data <- rbind(data,resamp)
table(rebal_schill_data$Schiller)[2]
table(rebal_schill_data$Schiller)[2]/dim(rebal_schill_data)[1]
```
```{r rebalacing Citology}
p = .20
records = as.integer(dim(data)[1])
rare = table(data$Citology)[2]
resamp_times <- (p*records - rare)/(1-p)


pos_cito <- which(data$Citology == "Positive")
resamp <- sample(pos_cito, size = resamp_times, replace = TRUE)
resamp <- data[resamp,]

rebal_cito_data <- rbind(data,resamp)
table(rebal_cito_data$Citology)[2]
table(rebal_cito_data$Citology)[2]/dim(rebal_cito_data)[1]
```

```{r rebalacing Biopsy}
p = .20
records = as.integer(dim(data)[1])
rare = table(data$Biopsy)[2]
resamp_times <- (p*records - rare)/(1-p)


pos_bio <- which(data$Biopsy == "Positive")
resamp <- sample(pos_bio, size = resamp_times, replace = TRUE)
resamp <- data[resamp,]

rebal_bio_data <- rbind(data,resamp)
table(rebal_bio_data$Biopsy)[2]
table(rebal_bio_data$Biopsy)[2]/dim(rebal_bio_data)[1]
```


```{r EDA, fig.height=14, fig.width=14, warning=FALSE}

# Histograms of data
library(Hmisc)
hist.data.frame(data[1:6], n.unique=1)
hist.data.frame(data[7:12], n.unique=1)
hist.data.frame(data[13:18], n.unique=1)
hist.data.frame(data[19:25], n.unique=1)
hist.data.frame(data[26:29], n.unique=1)

# correlation plot for the numeric variables
corrplot::corrplot(cor(data[,-c(5,10,12:19,21:29)]))

# correlation plot for the boolean variables
library(psych)
test <- data[,c(5,10,12:19,21:29)]
test <- test[,-c(15:18)]
test <- data.frame(lapply(test, as.numeric))
tetrachoric(test)
```
For the most part, random forest for classification has very similar properties to the regression analog discussed previously, including:  
* The model is relatively insensitive to values of mtry.  
* As with most trees, the data pre-processing requirements are minimal.  
* Out-of-bag measures of performance can be calculated, including accuracy,  sensitivity, specificity, and confusion matrices. 



```{r Logistic Regression Model}

cancer_predictors <- c("Age", "Number.of.sexual.partners", "First.sexual.intercourse", "Num.of.pregnancies", 
"Smokes", "STDs")

predictor_df <- subset(data, select = cancer_predictors)

logReg <- train(x = predictor_df, y = data$Schiller,
               method = "glm",
               preProcess = c("center","scale"))

#raw coefficients
data.frame(logReg$finalModel$coefficients)

#odds ratios
data.frame(exp(logReg$finalModel$coefficients))
```



```{r random forest model}



```

